---
title: "Partial-final-9jun"
author: "Yajie Guo"
date: "09/06/2025"
output: html_document
---

#generate the true dataset
```{r}

ssa.f.total<-function(theta,S0){
  
  Sini<-S0
  
  names(Sini)<-c('S1','S2')
  
  S.can<-S0
  
  t<-0
  
  Reaction<-matrix(c(-1,0,
                     0,+1,
                     0,-1),byrow=TRUE,ncol=2)  
  #Jump direction of (mRNA, protein) 
  
  
  # while(t[length(t)] <= Mt){
  while( !all(S.can == c(0, 0)) ){
    
    
    theta.can<-c(theta[1]*S.can[1], theta[2]*S.can[1], theta[3]*S.can[2]) 
    #calculate the current intensity function \lambda(x)=theta*x
    
    # Check if theta.can is c(0,0,0) and break the loop
    if (all(theta.can == c(0, 0, 0))) {
      break  # Exit the while loop
    }
    
    lambda<-sum(theta.can, na.rm = T) 
    
    wait<-rexp(1,rate=lambda) #waiting is exp(sum_intensities)
    
    
    t[length(t)+1] <-t[length(t)]+wait
    
    
    prob<-theta.can/lambda   
    
    R.can<-sample(size =1,x=c(1,2,3), prob = prob) #next jump direction
    
    
    
    if(R.can==1){
      S.can<-S.can+Reaction[1,] 
    }else{
      
      if(R.can==3){
        S.can<-S.can+Reaction[3,]
      }else{
        
        S.can<-S.can+Reaction[2,]
      }
    }
    
    
    Sini<-rbind.data.frame(Sini,S.can)
    
  }
  
  
  names(Sini)<-paste0('Species',c('mRNA','Protein'))
  
  jump.time.species<-cbind.data.frame(jump.times=t,Sini)
  return(jump.time.species)
}
```


```{r}
N=100 #scale
theta1=0.5 #true parameter
theta2.o= 90
theta3=3


S0 <- c(N, 0)
theta<-c(theta1,theta2.o/N,theta3)
result.out <- ssa.f.total(theta, S0)
```

#plot the true trajectory

```{r}
chop.T.index<- findInterval(10,result.out$jump.times)

range_y <- range( c(result.out$SpeciesmRNA[1:chop.T.index], result.out$SpeciesProtein[1:chop.T.index]))

range_x<- range(result.out$jump.times[1:chop.T.index])

plot(result.out$jump.times[1:chop.T.index], result.out$SpeciesProtein[1:chop.T.index],
     type = "n",                   # draw axes only
     ylim = range_y,
     xlim= range_x,
     xlab = "Time",
     ylab = "Molecule count",
     main = "Jump-process trajectories")

## ── 3. Add staircase lines for each species ─────────────────────
lines(result.out$jump.times[1:chop.T.index], result.out$SpeciesmRNA[1:chop.T.index],
      type = "s",  lwd = 2,col = "steelblue")        # mRNA: solid line

lines(result.out$jump.times[1:chop.T.index], result.out$SpeciesProtein[1:chop.T.index],
      type = "s",  lwd = 2, lty = 1,col = "firebrick")  # Protein: dashed line

## ── 4. Legend ───────────────────────────────────────────────────
legend("topright",
       legend = c("mRNA", "Protein"),
       lty    = c(1, 2),
       lwd    = 2,
       bty    = "n")

```


```{r}
#chop.T.index<- findInterval(20,result.out$jump.times)

range_y <- range(
  #c(result.out$SpeciesmRNA[1:chop.T.index],
  result.out$SpeciesProtein[1:chop.T.index])

#range_x<- range(result.out$jump.times[1:chop.T.index])

plot(result.out$jump.times[1:chop.T.index], result.out$SpeciesProtein[1:chop.T.index],
     type = "n",                   # draw axes only
     ylim = range_y,
     xlim= range_x,
     xlab = "Time",
     ylab = "Molecule count",
     main = "Jump-process trajectories")

## ── 3. Add staircase lines for each species ─────────────────────
#lines(result.out$jump.times[1:chop.T.index], result.out$SpeciesmRNA[1:chop.T.index],
#      type = "s",  lwd = 2,col = "steelblue")        # mRNA: solid line

lines(result.out$jump.times[1:chop.T.index], result.out$SpeciesProtein[1:chop.T.index],
      type = "s",  lwd = 2, lty = 1,col = "firebrick")  # Protein: dashed line

## ── 4. Legend ───────────────────────────────────────────────────
legend("topright",
       legend = c( "Protein"),
       lty    = c(2),
       lwd    = 2,
       bty    = "n")

 
```


#fn: given full trajectory, let us chop the [0,T] half and half, with chop.number = 2^{chop.number.time}

```{r}
step.fn<- function(all.data,chop.number.times){
   time.r<- all.data$jump.times
   
   cond <- all( all.data[(nrow(all.data)-1),2:3] == c(0,0) )   #if the species all die out before T, the reaction final is die out time
   
   if( cond ){
   reaction.final.time <-  time.r[length(time.r)-1]
   }else{  reaction.final.time<- time.r[length(time.r)]  }  #if the species not die out at time T, 
   
   
   chop.numebr= 2^{chop.number.times}
   
    each.inter.len <- reaction.final.time/(chop.numebr+1)   #interval number
    
   partial.equal.obse.time<- rep(NA,l=(chop.numebr+2))
   
   for(i in  1:chop.numebr){
     partial.equal.obse.time[i+1]<- each.inter.len * i
   }
   partial.equal.obse.time[1]<-time.r[1]
   partial.equal.obse.time[length(partial.equal.obse.time)]<-reaction.final.time
   
   return(list(partial.equal.obse.time=partial.equal.obse.time,chop.numebr=chop.numebr ))
}
```

#fn: chop time equally 


#know.reaction.final.index: we know [0,T], where T is from the real full trajectory data
#T.protein.unknown: if we do not know the true reaction stopped time, we need this variables to chop the full.data

```{r}
chop.time.equally.fn<- function( all.data, chop.number.times,  know.reaction.final.index, T.protein.unknown ){
  #if we known the whole reaction die out time, the [0,T] is the given by the last line of data
  if(know.reaction.final.index == TRUE){
    
    T.protein = all.data$jump.times[nrow(all.data)]
    with.T.all.data<- all.data
    
  }else{
    T.protein = T.protein.unknown
    #else we chop the time respect to some random observation time
    
    time.r<- all.data$jump.times
  if(time.r[length(time.r)]<= T.protein){
    index<- length(time.r)+1 #since the index is +1
    #if all false, the which.max will return 1
  }else{ index<- which.max( time.r >= T.protein)}
  
  
  all.data  <- all.data[1:(index-1),]
  last.interval<- as.data.frame( c(T.protein,  all.data [nrow( all.data ),-1] )  )
  colnames(last.interval)<- colnames(all.data)
  
  with.T.all.data <- rbind(all.data,last.interval)

  }
###################################################################################################
#give them jump events index
with.T.all.data$SpeciesProtein<- as.numeric(with.T.all.data$SpeciesProtein )
with.T.all.data$SpeciesmRNA <- as.numeric(with.T.all.data$SpeciesmRNA)
with.T.all.data$jump.type<- rep(NA, l= nrow(with.T.all.data))
for(j in 2:nrow(with.T.all.data) ){
  
  if ( with.T.all.data$SpeciesProtein[j] - with.T.all.data$SpeciesProtein[j-1] == +1 ){
    with.T.all.data$jump.type[j] <- c("birth.event")
    
  }
  if ( with.T.all.data$SpeciesmRNA[j] - with.T.all.data$SpeciesmRNA[j-1] == -1 ){
    with.T.all.data$jump.type[j]<- c("mrna.death.event")}
  if ( with.T.all.data$SpeciesProtein[j] - with.T.all.data$SpeciesProtein[j-1] == -1 ){
    with.T.all.data$jump.type[j]<- c("death.event")}
}
 
 
   with.T.all.data[,1:3]  <- data.frame(
  lapply(with.T.all.data[,1:3], function(col) as.numeric(col)),
  check.names = FALSE
)
###################################################################################################    
 #partial data
  all.out<- step.fn(all.data = with.T.all.data, chop.number.times = chop.number.times)
  
   partial.obse.times<- all.out$partial.equal.obse.time
   
   chop.number<- all.out$chop.numebr
   
   partial.all.species.data<- as.data.frame(matrix(NA, nrow= length(partial.obse.times), ncol=4))
    colnames( partial.all.species.data)<- c("partial.observed.time", "obs.SpeciesmRNA" ,   "obs.SpeciesProtein","jump.type")
    
   
   for(j in 1:length(partial.obse.times)){
     
  index <- findInterval(partial.obse.times[j], with.T.all.data$jump.times)
  one.row <-  cbind.data.frame("partial.observed.time"=partial.obse.times[j], 
                               "obs.SpeciesmRNA"= as.numeric( with.T.all.data[index,2]), 
                               "obs.SpeciesProtein"= as.numeric( with.T.all.data[index,3]),
                               "jump.type"="Partial.obse.times", deparse.level =0)
   partial.all.species.data[j,]<-  one.row 
   
  
   }
 
  
 return(list(partial.all.species.data=partial.all.species.data, 
             with.T.all.data= with.T.all.data, 
             T.protein= T.protein,  
             chop.number=  chop.number) )
}
```




```{r}
#T.protein = 1.5
data.p.out<- chop.time.equally.fn(all.data = result.out, chop.number.times=4,  know.reaction.final.index=TRUE, T.protein.unknown=NA )
data.p.out$partial.all.species.data  #partial observed data,[0,y_1,...,y_{chop.number}, T]
data.p.out$with.T.all.data  #full-observed data during [0,T]
data.p.out$chop.number #this the chop number 
```


#PARAMETER MCMC START: 

#STEP1: GIBBS of PARAMETERs---- FULL CONDITIONAL GAMMA DISTRIBUTION 

#Attention:
#each interval data is updated term by term, but the evaluation here is base on the full trajectory!
 
#fn:summary of data to input posterior gamma function
```{r}
#all.data.T will update 
cond.para.based.full.data.fn<- function(all.data.T){
  
#extract the components in ith interval 
time.r <-   as.numeric( all.data.T$jump.times  )  #include the head and tail of the partial observed data
protein.full <-  as.numeric(  all.data.T$SpeciesProtein  )
mrna.full <- as.numeric( all.data.T$SpeciesmRNA  )
#extract the na,nb,nd and integral of X1,X2  
       l.t<- length(time.r) #total is with the head and tail T
       int.X2<- 0
       int.X1<-0
       nd =0  #protein death events number
       na =  mrna.full[1] - mrna.full[length(mrna.full)]  #mRNA degradation number
       nb=0
       for(i in 2:l.t){
         
          int.X2<- int.X2+ (time.r[i]-time.r[i-1])*protein.full[i-1]
          
          int.X1<- int.X1 +(time.r[i]-time.r[i-1])*mrna.full[i-1]
          
      if(protein.full[i]-protein.full[i-1] == -1){  nd = nd +1}
          if( protein.full[i]-protein.full[i-1] == +1){nb=nb+1}
          
          }
  
       #nb= (l.t-2)- na - nd #protein birth events number

  return(list(int.X1=int.X1,int.X2=int.X2, na=na,nb=nb,nd=nd))     
}  
```

```{r}
data.T.summary.out<- cond.para.based.full.data.fn(all.data.T = data.p.out$with.T.all.data )
data.T.summary.out
```




#fn: three posterior gamma function


```{r}
full.cond.theta.gamma.fn<- function(full.data.summary.out.T, N,alp1, alp2, bet1, bet2,alp3, bet3 ){
  
  na= full.data.summary.out.T$na
nb= full.data.summary.out.T$nb
nd= full.data.summary.out.T$nd
int.X1= full.data.summary.out.T$int.X1
int.X2= full.data.summary.out.T$int.X2

#return scaled version theta2
  a1<- na + alp1
  b1<-int.X1  +bet1
  
  the1.can<-rgamma(1, shape=a1,rate=b1)
  
  a2<- nb +alp2
  b2<-  int.X1/N +   bet2 
  #here the theta2 is the the scaled version, like 10/n, our theta2.out is 10
  the2.can<-rgamma(1,shape=a2,rate=b2)
  
  a3=nd  +alp3
  b3=int.X2  +bet3 
  the3.out <- rgamma(n=1,shape=a3 , rate=b3 )
  
  theta.all.can<- c( the1.can, the2.can,the3.out )
 return(theta.all.can)  
 }
 
```

```{r}
 alp1=0.0001
  alp2=0.0001
  bet1=0.0001
  bet2=0.0001
  alp3=0.0001
  bet3= 0.0001
  
full.cond.theta.gamma.fn(full.data.summary.out.T = data.T.summary.out,
                        alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3, N=N)

```






 
#STEP2: GIBBS of mRNA---- FULL CONDITIONAL truncated exponential distribution 





 
#fn: each partial interval updates of mrna

```{r}

ith.interval.mrna.samples.fixed.jumps.fn<- function(ith.all.data.T, theta,N){
  theta1<- theta[1]
  theta2.o<- theta[2]
    a<- theta1+ theta2.o/N  #this is the parameters needed to be updated afterwards
    
  protein.born.index<- which( ith.all.data.T$jump.type == "birth.event" )
  protein.birth.times<- as.numeric( ith.all.data.T$jump.times[protein.born.index] )
  
  mrna.ith.jumps.index<- which( ith.all.data.T$jump.type == "mrna.death.event")
  
  mrna.ith.all.states  <- rbind( ith.all.data.T[1,], 
                                 ith.all.data.T[mrna.ith.jumps.index,], 
                                 ith.all.data.T[nrow(ith.all.data.T),] ) #with head and tail
  
  
  # yi.update.len = length(mrna.ith.jumps.index)+1
  yi.update.len = length(mrna.ith.jumps.index)+1
  if(yi.update.len==1){yi.update.len=2}
  
  for(i in 2:yi.update.len) {
    
    yi.minus.1 <- mrna.ith.all.states$jump.times[i-1]
    yi.plus.1 <-  mrna.ith.all.states$jump.times[i+1]
    
    
    
    which.protein.index <- protein.birth.times > yi.minus.1 & protein.birth.times <  yi.plus.1
    
    Region.i.protein.born <- protein.birth.times[which.protein.index]
    #now the region becomes [ yi.minus.1, #Region.i.protein.bron, yi.plus.1 ]  
    
    Region.i.protein.bron.number<- length(Region.i.protein.born)
    #length of subintervals is Region.i.protein.bron.number+1
    
    sub.interval.yi<-c(  yi.minus.1,Region.i.protein.born, yi.plus.1)
    
    
    #before the yi the jumps, with the X_1(y_{i-1}) weight
    weight.i.minus.mRNA <-     mrna.ith.all.states$SpeciesmRNA[i-1]
    
    #after the yi the jumps, with the X_1(y_{i-1}) weight
    weight.i.mRNA<-  mrna.ith.all.states$SpeciesmRNA[i]
    
   
    
    if(weight.i.mRNA == 0){
      #the last death event of mrna molecule need to later than protein bith events
      ratio.sub.interval<- rep(0,l=(Region.i.protein.bron.number+1) )
      ratio.sub.interval[length(ratio.sub.interval)]<- 1
      
    }else{
      ################################################################################   
      #Each subinterval have each weight
      log.total.weight.vec<-rep(NA,l=(Region.i.protein.bron.number+1) )
      
      log.subinterval.size.vec<- rep(NA,l=(Region.i.protein.bron.number+1) )
       
      #calculate the sub-interval size 
      #here we normalized the weight
       
      #####################################################################################################      
      
      for(k in 1:(Region.i.protein.bron.number+1)){
        
        # total.weight.vec[k]<- (weight.i.minus.mRNA^(k-1))*( weight.i.mRNA^(Region.i.protein.bron.number-(k-1)) )  
        
        log.total.weight.vec[k]<- (k-1)*log(weight.i.minus.mRNA)+ (Region.i.protein.bron.number-(k-1))*log( weight.i.mRNA )  
        
        log.subinterval.size.vec[k]<- log.total.weight.vec[k]+ log((exp(-a*sub.interval.yi[k])-exp(-a*sub.interval.yi[k+1]))*a^{-1})
        
        # if(is.na(log.subinterval.size.vec[k]) ){ print(log.total.weight.vec[k]) }
        
      }
      
      max.log.size<- max(log.subinterval.size.vec)
      
      log_sum_all_interval_size <- max.log.size + 
        log( sum( exp(log.subinterval.size.vec -max.log.size )) )
      
      # ratio.sub.interval<- subinterval.size.vec/sum(subinterval.size.vec)
      
      #if(is.na(log_sum_all_interval_size) ){ print(max.log.size) }
      
      
      ratio.sub.interval<- exp( log.subinterval.size.vec - log_sum_all_interval_size)
      #####################################################################################################   
    }     
    
    
    
    #update into this region
    index.yi.updated.region<- sample(x=(Region.i.protein.bron.number+1),size=1, prob=ratio.sub.interval)
    
    pos.updated.yi.region<- sub.interval.yi[index.yi.updated.region:(index.yi.updated.region+1)]
    
    #inverse sampling  
    u=runif(n=1)
    #this is new yi
    yi.can<- -a^{-1}*log( exp(-a*pos.updated.yi.region[1])-u*(exp(-a*pos.updated.yi.region[1])-exp(-a*pos.updated.yi.region[2]) ) )
    
    #put it back to the original ith interval data
    mrna.ith.all.states$jump.times[i]<- yi.can #where the mrna and protein states unchange
    
  }
  
  #update back to the ith interval data
  
  ith.all.data.T.del.mrna <- ith.all.data.T[-c(mrna.ith.jumps.index),]
  
  
  for(j in 2:(nrow(mrna.ith.all.states)-1) ){
    
    new.mrna.jumps <- mrna.ith.all.states$jump.times[j]    #without head and tail,just mrna jumps
    
    index.mrna<- findInterval( new.mrna.jumps,  as.numeric(ith.all.data.T.del.mrna$jump.times))
    
    ith.all.data.T.del.mrna <-  rbind(ith.all.data.T.del.mrna[1:index.mrna,],
                                      c( new.mrna.jumps, NA, NA, "mrna.death.event"),
                                      ith.all.data.T.del.mrna[(index.mrna+1):nrow(ith.all.data.T.del.mrna),]
    )
    
  }
  re.ith.all.data.T<- all.species.reorder.fn(ith.all.data.T= ith.all.data.T.del.mrna )
  
  return(re.ith.all.data.T)
  
}

```

#fn: take ith interval of the fullly observed data out 

```{r}
extract.ith.data.T.fn <-function(partial.data.out, all.data.T, k){
  
  y.k1<- partial.data.out$partial.observed.time[k]
  y.k<-  partial.data.out$partial.observed.time[k+1]
  
error<- 1e-12

  row.head <- cbind(partial.data.out[k,])
 row.tail<- cbind(partial.data.out[(k+1),])

 index.chop.all.data<-   which( all.data.T$jump.times > y.k1+error  & all.data.T$jump.times < y.k-error   )
   
 if( length(index.chop.all.data) == 0){
   ith.all.data.T <- rbind(  row.head,
                            row.tail, deparse.level = 0 )
 }else{
  
 chop.ith.all.data.T <- all.data.T[index.chop.all.data[1]:index.chop.all.data[length(index.chop.all.data)], ]
 #this the ith interval data
  colnames( row.head)<- colnames(all.data.T)
  colnames( row.tail)<- colnames(all.data.T)
  
 ith.all.data.T <- rbind(  row.head,
                         chop.ith.all.data.T,
                         row.tail, deparse.level = 1 )
 
 }
 colnames(ith.all.data.T)<- colnames(all.data.T)
 
  ith.all.data.T[,1:3]  <- data.frame(
  lapply(ith.all.data.T[,1:3], function(col) as.numeric(col)),
  check.names = FALSE
)
  
 return(ith.all.data.T)
 }
```

```{r}
extract.ith.data.T.fn(partial.data.out = data.p.out$partial.all.species.data, all.data.T= data.p.out$with.T.all.data, k=1 )
```

#fn: update whole mrna states

```{r}
ful.cond.mrna.jumps.updated.fixed.na.fn<- function( theta, partial.data.out, all.data.T,N){
  
   all.data.T[,1:3]  <- data.frame(
  lapply(all.data.T[,1:3], function(col) as.numeric(col)),
  check.names = FALSE
)
  theta1<- theta[1]
  theta2.o<- theta[2]
  theta3<- theta[3]
chop.number<- length(partial.data.out[,1])-2
updated.all.data.T<- all.data.T[1,]
   
  #for each interval protein update
##################################################################################################
for(k in 1:(chop.number+1) ){
  
  #extract the all data within the ith partial time observation time 
 ith.all.data.T<- extract.ith.data.T.fn(partial.data.out = partial.data.out, all.data.T=  all.data.T, k=k )
 #give the ith data
 ##################################################################################################
 
 #no mrna jumps, no mran samples
 if( !any( ith.all.data.T$jump.type== c("mrna.death.event") ) ){
 mrna.updated.ith.all.data.T<- ith.all.data.T
 }else{
 #let us update all the ith mran jump times during each partial time observation
mrna.updated.ith.all.data.T<-  ith.interval.mrna.samples.fixed.jumps.fn(ith.all.data.T =ith.all.data.T, theta = theta, N=N )
 }
 #print(k)
 
 colnames(mrna.updated.ith.all.data.T)<- colnames(ith.all.data.T)
 
   updated.all.data.T<- rbind(updated.all.data.T, mrna.updated.ith.all.data.T[-1, ])
 
}

return(updated.all.data.T)
}
```





 
#STEP3: MH of PROTEIN---- Propose the three types of moves(adding, deleting, or shifting)

#protein MH updates 

#fn: after some moves of events, reorder of all species

```{r}

all.species.reorder.fn<-function(ith.all.data.T){
  
  #reorder the species molecule number
  ith.all.data.T$SpeciesmRNA<- as.numeric(ith.all.data.T$SpeciesmRNA)
  ith.all.data.T$SpeciesProtein<- as.numeric(ith.all.data.T$SpeciesProtein)
  
  if( all(ith.all.data.T$jump.type == "Partial.obse.times")){
     return(ith.all.data.T)
  }
  
  for(l in 2:(nrow(ith.all.data.T)-1)){
    
    if ( ith.all.data.T$jump.type[l] == "mrna.death.event" ) {
      
      ith.all.data.T$SpeciesProtein[l]<- ith.all.data.T$SpeciesProtein[l-1]
      ith.all.data.T$SpeciesmRNA[l]<- ith.all.data.T$SpeciesmRNA[l-1]-1
      
    }else{
      
      if( ith.all.data.T$jump.type[l] == "birth.event"){
        ith.all.data.T$SpeciesProtein[l]<-  ith.all.data.T$SpeciesProtein[l-1] +1 
        ith.all.data.T$SpeciesmRNA[l]<- ith.all.data.T$SpeciesmRNA[l-1]
      } 
      if(ith.all.data.T$jump.type[l] == "death.event" ){
        ith.all.data.T$SpeciesProtein[l]<-  ith.all.data.T$SpeciesProtein[l-1]  - 1 
        ith.all.data.T$SpeciesmRNA[l]<- ith.all.data.T$SpeciesmRNA[l-1]
      }
    }
  }
  ith.all.data.T[,1:3]  <- data.frame(
    lapply(ith.all.data.T[,1:3], function(col) as.numeric(col)),
    check.names = FALSE
  )
  
  return(ith.all.data.T)
  
}


```




#Protein events
#fn1: adding a pair
```{r}

protein.adding.pair.fn<- function(ith.all.data.T){

  #new birth and death data  
 yi.minus.1<- ith.all.data.T$jump.times[1]
  yi<-ith.all.data.T$jump.times[nrow(ith.all.data.T)] 
 
  add.sams <-  runif(n=2, min=yi.minus.1, max=yi)
  birth.time.add <- add.sams[1]
  death.time.add<- add.sams[2]
   
  
 birth.time.index <-   findInterval( birth.time.add, as.numeric( ith.all.data.T$jump.times)  )  # the ith interavl is index by [i-1,i]
 
   
ith.all.data.T  <- rbind(ith.all.data.T[1:birth.time.index, ], 
  c(birth.time.add, NA, NA, "birth.event"),   
        ith.all.data.T[(birth.time.index+1):nrow(ith.all.data.T), ])

#this need to be later, where the ith.all.data.T is reordered!

   death.time.index <-   findInterval( death.time.add, as.numeric( ith.all.data.T$jump.times)  )
 
ith.all.data.T<- rbind(ith.all.data.T[1:death.time.index, ], 
                   c(death.time.add, NA, NA, "death.event"),   ith.all.data.T[(death.time.index+1):nrow(ith.all.data.T), ])
 
re.ith.all.data.T<- all.species.reorder.fn(ith.all.data.T=ith.all.data.T)
return( re.ith.all.data.T )
}
```

#fn2: deleting a pair
```{r}
protein.deleting.pair.fn<- function(ith.all.data.T ){
  
  
  birth.index <-  which(ith.all.data.T$jump.type == "birth.event")
  death.index <-  which(ith.all.data.T$jump.type == "death.event") 
  
  
  if( length(birth.index)>1 ){
 birth.delete.sam <-  sample( birth.index, size=1)}else{  birth.delete.sam <-  birth.index }
  
  if( length(death.index) >1 ){
  death.delete.sam  <- sample(death.index, size=1)
  }else{death.delete.sam  <-  death.index  }
  
  
 ith.all.data.T<- ith.all.data.T[-c( birth.delete.sam, death.delete.sam), ]
 
 
 re.ith.all.data.T<- all.species.reorder.fn(ith.all.data.T=ith.all.data.T)
 
 return(re.ith.all.data.T)
 
}
```

#fn3: shifting one of the events
```{r}
protein.shift.pair.fn<- function( ith.all.data.T){
  
 
index.all.events <- which( ith.all.data.T$jump.type %in% c("birth.event", "death.event"))

if(length(index.all.events)==1){
  shift.index<- index.all.events
}else{
  shift.index <- sample(c(index.all.events), size = 1)
}

ith.all.data.T$jump.times<- as.numeric(ith.all.data.T$jump.times)

 r.t<- runif(n=1,min=ith.all.data.T$jump.times[1],
              max= ith.all.data.T$jump.times[nrow(ith.all.data.T)])
  


 shifted.row <-  ith.all.data.T[shift.index, ] #keep this shifted row information
  shifted.row$jump.times = r.t
  shifted.row$SpeciesmRNA = NA
  
  ith.all.data.T<- ith.all.data.T[-shift.index, ]
  
  added.index<- findInterval(r.t, ith.all.data.T$jump.times    )
    
 ith.all.data.T  <- rbind(ith.all.data.T[1:added.index, ], 
                     shifted.row, #added the previous shifted one
        ith.all.data.T[(added.index+1):nrow(ith.all.data.T), ])
 
 
#reorder the protein value
 
 re.ith.all.data.T<- all.species.reorder.fn(ith.all.data.T=ith.all.data.T)
 
 return(re.ith.all.data.T)

}
```


#fn: accept or not basing on this ratio of ith interval density

```{r}
##########################################################################################################
log.density.protein.cond.mrna.jumps.fn<- function(ith.all.data.T,N,theta){
  
  if( !all(ith.all.data.T$SpeciesProtein >=0) ){
    data2.cond.data1= -Inf 
  return( data2.cond.data1 )
  }
  
  
  
theta1= theta[1]
theta2.o= theta[2]
theta3= theta[3]
 #############################################################################################
  ith.protein.d <-  as.numeric( ith.all.data.T$SpeciesProtein )
  ith.mrna.d <- as.numeric( ith.all.data.T$SpeciesmRNA )
  ith.times<- as.numeric( ith.all.data.T$jump.times )
  int.X1.ith<- 0 
  int.X2.ith<- 0 
  
log.prod.x1<- 0  #this may not needed, because mran does not change during the ith interval
 log.prod.x2<-0
 
 
  for(j in 2:length(ith.times)){
    int.X1.ith<- int.X1.ith +   ith.mrna.d[j-1]*( ith.times[j]- ith.times[j-1] )
    int.X2.ith<- int.X2.ith + ith.protein.d[j-1]*(ith.times[j]- ith.times[j-1] )
    
    if(ith.protein.d[j]- ith.protein.d[j-1] == +1 ){
    log.prod.x1 <-  log.prod.x1 + log( theta2.o*ith.mrna.d[j]/N )}
    
    if(ith.protein.d[j]- ith.protein.d[j-1] == -1){
      log.prod.x2<- log.prod.x2 + log(theta3*ith.protein.d[j-1] )
    }
  }
 
  if(theta1>0 &&  theta2.o>0 && theta3>0 ){
    
 term1<-  log.prod.x1 - ( theta2.o/N )*int.X1.ith 
term2<- log.prod.x2 - theta3*int.X2.ith  
    
  data2.cond.data1<- term1+term2
  
  }else{  data2.cond.data1=-Inf }
  return( data2.cond.data1 )
}
```


#states of protein move type


#fn: case1 moves for current species states(0,0)


```{r}
case1.00.move.protein.fn<- function(ith.all.data.T, weight.prob ){
  
  new.ith.all.data.T<- protein.adding.pair.fn(ith.all.data.T = ith.all.data.T)
   
 #calculate the weight proposal ratio  
##########################################################  
    #when nb=bd=0, only adding prob.weig=1,others are 0, the reverse are not, three possibles, with weight p_2 of deleting
   
   yi.min.1<- ith.all.data.T$jump.times[1]
   yi<- ith.all.data.T$jump.times[nrow(ith.all.data.T)]
   
   proporsal.weight.ratio<- log( weight.prob[2]) + 2*log( yi - yi.min.1 ) 
   
   move.type<- "add_a_pair"
   
   return(list(new.ith.all.data.T=new.ith.all.data.T , proporsal.weight.ratio= proporsal.weight.ratio, move.type=move.type)) 
}
```

#fn: case2 moves for current species states(0,>=1) or (>=1,0)

```{r}
case2.one.jump.type.protein<-function(ith.all.data.T, weight.prob){
    
  norm.prob<-c(weight.prob[1], weight.prob[3])/(weight.prob[1]+weight.prob[3])
  
    u.2<- sample(c(1,2), size=1, prob =norm.prob) #equal of add or shift
  
       yi.min.1<- ith.all.data.T$jump.times[1]
   yi<- ith.all.data.T$jump.times[nrow(ith.all.data.T)]
   
   
    if(u.2 == 1){
       new.ith.all.data.T<-   protein.adding.pair.fn(ith.all.data.T = ith.all.data.T)
      
  #calculate the weight proposal ratio  
##########################################################    
      new.nb <- length(which( new.ith.all.data.T$jump.type == "birth.event"))
     new.nd <- length(which( new.ith.all.data.T$jump.type == "death.event"))
     
 proporsal.weight.ratio<- log( weight.prob[2]/norm.prob[1]) + 2*log( yi - yi.min.1 )- log(new.nb*new.nd )
          move.type<- "add_a_pair"
          
 ##########################################################       
    }else{
        
       new.ith.all.data.T<- protein.shift.pair.fn( ith.all.data.T = ith.all.data.T)
#calculate the weight proposal ratio  
##########################################################     
        proporsal.weight.ratio<- log(1)
##########################################################
         move.type<- "shift_a_event"
    }
   
   
    return(list(new.ith.all.data.T= new.ith.all.data.T, proporsal.weight.ratio= proporsal.weight.ratio,  move.type=  move.type)) 
   
}
```


#fn: case3 moves for current species states(>=1,>=1)

```{r}
case3.ge1move.protein.fn<- function(ith.all.data.T, weight.prob ){
    
  three.rv <- sample(c(1,2,3), size=1, prob= weight.prob)
   norm.prob= c(weight.prob[1], weight.prob[3])/(weight.prob[1]+weight.prob[3])
     
   yi.min.1<- ith.all.data.T$jump.times[1]
   yi<- ith.all.data.T$jump.times[nrow(ith.all.data.T)]
   
   #adding 
  if( three.rv == 1 ){
   # print("add a pair.by.three.possibles")
    
  new.ith.all.data.T<- protein.adding.pair.fn(ith.all.data.T = ith.all.data.T)
      new.nb <- length(which( new.ith.all.data.T$jump.type == "birth.event"))
     new.nd <- length(which( new.ith.all.data.T$jump.type == "death.event"))
     
 #calculate the weight proposal ratio  
##########################################################    
   proporsal.weight.ratio<- log( weight.prob[2]/weight.prob[1]) + 2*log(yi -yi.min.1)- log(new.nb*new.nd )}
 ##########################################################     
      move.type<- "add_a_pair"
 
   #shifting      
if( three.rv == 3 ){
 
    #print("shift a event.by.three.possibles")
 new.ith.all.data.T<- protein.shift.pair.fn( ith.all.data.T = ith.all.data.T)
#calculate the weight proposal ratio  
##########################################################  
 proporsal.weight.ratio<- log(1)
 ########################################################## 
  move.type<- "shift_a_event"
}

    #deleting   
  if( three.rv == 2 ){
   
    # print("delete a pair.by.three.possibles")
  new.ith.all.data.T<- protein.deleting.pair.fn( ith.all.data.T = ith.all.data.T) 

  #calculate the weight proposal ratio  
  ##########################################################
      old.nb <- length(which( ith.all.data.T$jump.type == "birth.event"))
     old.nd <- length(which( ith.all.data.T$jump.type == "death.event"))
     move.type<- "delete_a_pair"
     
  if( old.nb== 1 && old.nd ==1){
    norm.prob.ratio <-  log(1/weight.prob[2])
  }else{
     if( (old.nb >=2 && old.nd==1 ) || (old.nb == 1 && old.nd>= 2 )  ){
        norm.prob.ratio <- log(norm.prob[1]/weight.prob[2])
     }
     
     if( old.nb>= 2 && old.nd >=2 ){
       norm.prob.ratio<- log(weight.prob[1]/ weight.prob[2])
     }
    
  }
  
  proporsal.weight.ratio<-  norm.prob.ratio +  log(old.nb*old.nd)- 2*log(yi -yi.min.1)
   
   }
       
    
   return(list(new.ith.all.data.T= new.ith.all.data.T, proporsal.weight.ratio= proporsal.weight.ratio, move.type=move.type)) 
}
```




#fn: update one of the interval term by term and return the full trajectory of protien

```{r}
#events.move<- list("add"= 0, "delete"=0,"shift"=0)

protein.all.data.update.fn <- function(all.data.T, partial.data.out, weight.prob, theta, N){
   
updated.all.data.T <- all.data.T[1,] #store all the updated data
 # norm.prob= c(weight.prob[1], weight.prob[3])/(weight.prob[1]+weight.prob[3]) 
chop.number<- length(partial.data.out[,1])-2
  #for each interval protein update
 count=0
 
  interval.protein.birth.number<- rep(NA,l=(nrow(partial.data.out)-1) )
  
  
   events.move <- as.data.frame( matrix(0, nrow= 1, ncol=3) )
  colnames(events.move)<- c("add","shift","delete")
  
##################################################################################################
for(k in 1:(chop.number+1) ){
  
  #kth interval is from the index of k:k+1
ith.all.data.T<- extract.ith.data.T.fn(partial.data.out = partial.data.out, k=k, all.data.T = all.data.T)
  
 #let us start the three protein move types: ####################################################################################################################################################################################################
#case 1:(0,0)
cond1<- all(ith.all.data.T$jump.type %in% c("Partial.obse.times", "mrna.death.event") )
#case 3: (>=1,>=1), there are three possible moves
cond3<- all( c("birth.event", "death.event") %in% ith.all.data.T$jump.type ) 
#case2: their complements
cond2<- !(cond1 |cond3)


 #this the all data within the kth interval of partially observed data
  #if there is no events of protein, only add a pair is possible
  if(cond1  ){
   case1.out<- case1.00.move.protein.fn(ith.all.data.T= ith.all.data.T, weight.prob=weight.prob)
   new.ith.all.data.T<-  case1.out$new.ith.all.data.T
   new.proporsal.weight.ratio<- case1.out$proporsal.weight.ratio
   propose.type<- case1.out$move.type
##########################################################
  } 

 #case 2: (0,>=1) or (>=1, 0)  
  #if there is no death/birth events of protein during this interval, only adding or shift weight equally
 ##############################################################################################################

if(cond2){
 case2.out<-  case2.one.jump.type.protein(ith.all.data.T= ith.all.data.T, weight.prob=weight.prob)
  new.ith.all.data.T<-  case2.out$new.ith.all.data.T
   new.proporsal.weight.ratio<- case2.out$proporsal.weight.ratio
    propose.type<- case2.out$move.type
} 


#########################################################################################################
 if( cond3 ){
   case3.out<- case3.ge1move.protein.fn(ith.all.data.T= ith.all.data.T, weight.prob=weight.prob)
 new.ith.all.data.T<-  case3.out$new.ith.all.data.T
   new.proporsal.weight.ratio<- case3.out$proporsal.weight.ratio
    propose.type<- case3.out$move.type
   }


#finished cases

 #finished the proposal move, let us decide to accept it or not
##################################################################################################
##################################################################################################
#MH of protein update
 #old ith density
 #input mrna data into the density of protein birth events
old.d2.cond.d1.density <- log.density.protein.cond.mrna.jumps.fn( ith.all.data.T = ith.all.data.T, theta= theta, N=N)
 
 #new ith density
new.d2.cond.d1.density <- log.density.protein.cond.mrna.jumps.fn( ith.all.data.T = new.ith.all.data.T, theta= theta, N=N)
 
 density.log.ratio <-  new.d2.cond.d1.density-  old.d2.cond.d1.density
    
##################################################################################################

 u.rv<- log( runif(n=1) )
 
 alpha.i<-  density.log.ratio +  new.proporsal.weight.ratio  
 
 if(u.rv < alpha.i ){
   current.ith<- new.ith.all.data.T
   #print(paste("accept the new", k,"th interval data"))
   count= count+1
   
   if( propose.type ==  "add_a_pair"){ events.move$add<- events.move$add +1 }
   if(propose.type == "shift_a_event" ){events.move$shift <-events.move$shift +1 }
   if(propose.type == "delete_a_pair"  ){ events.move$delete <- events.move$delete +1  }
   
   
 }else{
   # print(paste("Not accept the new", k,"th interval data"))
    current.ith<-ith.all.data.T
 }
 
 
##################################################################################################
##################################################################################################
 interval.protein.birth.number[k]<- length(which(current.ith$jump.type == "birth.event"))
##################################################################################################

updated.all.data.T<- rbind(updated.all.data.T, current.ith[2:nrow(current.ith), ] )

}

  #finshed ith interval updates
  acceptance.protein<- count/(chop.number+1)
 # print(paste("accpetance rate of protein for one realization is", acceptance.protein))
  
  #this include the partial observed time
  return(list(updated.all.data.T=updated.all.data.T, 
                 acceptance.protein= acceptance.protein,
              interval.protein.birth.number=interval.protein.birth.number,
              events.move= events.move ))
}

```
 
 


```{r}
protein.all.data.update.simple.para.acceptance.fn<- function(all.data.T, partial.data.out, weight.prob, theta, N){
   
updated.all.data.T <- all.data.T[1,] #store all the updated data
  
chop.number<- length(partial.data.out[,1])-2
  #for each interval protein update
 count=0
 
 # interval.protein.birth.number<- rep(NA,l=(nrow(partial.data.out)-1) )
  
  
   events.move <- as.data.frame( matrix(0, nrow= 1, ncol=3) )
  colnames(events.move)<- c("add","shift","delete")
  
##################################################################################################
for(k in 1:(chop.number+1) ){
  
  #kth interval is from the index of k:k+1
ith.all.data.T<- extract.ith.data.T.fn(partial.data.out = partial.data.out, k=k, all.data.T = all.data.T)
  
 #let us start the three protein move types: #####################################################################################################################################################
#case 1:(0,0)
cond1<- all(ith.all.data.T$jump.type %in% c("Partial.obse.times", "mrna.death.event") )
#case 3: (>=1,>=1), there are three possible moves
cond3<- all( c("birth.event", "death.event") %in% ith.all.data.T$jump.type ) 
#case2: their complements
cond2<- !(cond1 |cond3)


 #this the all data within the kth interval of partially observed data
  #if there is no events of protein, only add a pair is possible
  if(cond1  ){
   case1.out<- case1.00.move.protein.fn(ith.all.data.T= ith.all.data.T, weight.prob=weight.prob)
   new.ith.all.data.T<-  case1.out$new.ith.all.data.T
   new.proporsal.weight.ratio<- case1.out$proporsal.weight.ratio
   propose.type<- case1.out$move.type
##########################################################
  } 

 #case 2: (0,>=1) or (>=1, 0)  
  #if there is no death/birth events of protein during this interval, only adding or shift weight equally
 ##############################################################################################################

if(cond2){
 case2.out<-  case2.one.jump.type.protein(ith.all.data.T= ith.all.data.T, weight.prob=weight.prob)
  new.ith.all.data.T<-  case2.out$new.ith.all.data.T
   new.proporsal.weight.ratio<- case2.out$proporsal.weight.ratio
    propose.type<- case2.out$move.type
} 


#########################################################################################################
 if( cond3 ){
   case3.out<- case3.ge1move.protein.fn(ith.all.data.T= ith.all.data.T, weight.prob=weight.prob)
 new.ith.all.data.T<-  case3.out$new.ith.all.data.T
   new.proporsal.weight.ratio<- case3.out$proporsal.weight.ratio
    propose.type<- case3.out$move.type
   }


#finished cases

 #finished the proposal move, let us decide to accept it or not
##################################################################################################
##################################################################################################
#MH of protein update
 #old ith density
 #input mrna data into the density of protein birth events
old.d2.cond.d1.density <- log.density.protein.cond.mrna.jumps.fn( ith.all.data.T = ith.all.data.T, theta= theta, N=N)
 
 #new ith density
new.d2.cond.d1.density <- log.density.protein.cond.mrna.jumps.fn( ith.all.data.T = new.ith.all.data.T, theta= theta, N=N)
 
 density.log.ratio <-  new.d2.cond.d1.density-  old.d2.cond.d1.density
    
##################################################################################################

 u.rv<- log( runif(n=1) )
 
 alpha.i<-  density.log.ratio +  new.proporsal.weight.ratio  
 
 if(u.rv < alpha.i ){
   current.ith<- new.ith.all.data.T
   #print(paste("accept the new", k,"th interval data"))
   count= count+1
   
   if( propose.type ==  "add_a_pair"){ events.move$add<- events.move$add +1 }
   if(propose.type == "shift_a_event" ){events.move$shift <-events.move$shift +1 }
   if(propose.type == "delete_a_pair"  ){ events.move$delete <- events.move$delete +1  }
   
   
 }else{
   # print(paste("Not accept the new", k,"th interval data"))
    current.ith<-ith.all.data.T
 }
 
 
##################################################################################################
##################################################################################################
# interval.protein.birth.number[k]<- length(which(current.ith$jump.type == "birth.event"))
##################################################################################################

updated.all.data.T<- rbind(updated.all.data.T, current.ith[2:nrow(current.ith), ] )

}

  #finshed ith interval updates
  acceptance.protein<- count/(chop.number+1)
 # print(paste("accpetance rate of protein for one realization is", acceptance.protein))
  
  #this include the partial observed time
  return(list(updated.all.data.T=updated.all.data.T, 
                 acceptance.protein= acceptance.protein,
             # interval.protein.birth.number=interval.protein.birth.number,
              events.move= events.move ))
}
```




```{r}

one.mcmc.run.fn<- function(all.data.T, partial.data.out, weight.prob, para.iter, initial.para.states, N, alp1,alp2, alp3, bet1, bet2,bet3 ){
  
    #input the scaled theta, where theta2 is not theta2/N here
  
  one.para.out<- matrix(NA, nrow= (para.iter+1), ncol=3)
  one.para.out[1,]<- initial.para.states
  
  total.events.move.out<- as.data.frame( matrix(NA, nrow= para.iter, ncol=3) )
  colnames(total.events.move.out)<- c("add","shift","delete")
  
  #start each data, one row of parameter corresponding to one of data matrix
  store.all.data.T<- vector("list", (para.iter+1))
  store.all.data.T[[1]]<- all.data.T
  
  #acceptance.rate.protein<- rep(NA,l=para.iter)
  protein.birth.number<- rep(NA,l=para.iter)
  #start from true protein birth number
  protein.birth.number[1]<- length(which(all.data.T$jump.type == "birth.event"))
  
  total.interval.protein.birth.number<- matrix(NA, nrow=para.iter, ncol= (nrow(partial.data.out)-1) )
  
  for(i in 2:(para.iter+1)){
#step 1: Gibbs of paramters 
#conditional on the [0,T] of full-observed data
#gibbs update: which is consider the full trajectory not in the ith interval!
 ##################################################################################################
 #this is the info for the summary of data to input the  the full-conditional of parameter 
full.data.summary.out.T.value<- cond.para.based.full.data.fn(all.data.T = all.data.T)
  

#this is the function of full conditional of parameter out given the full conditional data
  theta.all.can<- full.cond.theta.gamma.fn(full.data.summary.out.T= full.data.summary.out.T.value,
                        alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3, N=N)
  
  
   one.para.out[i,]<- theta.all.can
##################################################################################################
  
  
#step 2: Gibbs of mRNA
#conditional on the [0,T] of protein birth data & N jumps of the total reaction times of mRNA data 
#because the mRNA jumps is conditional on the protein birth events
#2.1 input of protein birth data(X2.data)
##################################################################################################
   
mrna.upated.all.data.T <- ful.cond.mrna.jumps.updated.fixed.na.fn(theta= theta.all.can, partial.data.out = partial.data.out, all.data.T = all.data.T,  N=N)
##################################################################################################
#step3: MH of protein  
#this is the update of all intervals 
##################################################################################################
   
 
protein.all.data.T <- protein.all.data.update.fn(all.data.T= mrna.upated.all.data.T, partial.data.out=partial.data.out, weight.prob=weight.prob, theta=theta.all.can, N=N )
  
   
   all.data.T<-  protein.all.data.T$updated.all.data.T  #update the current species data
   
   total.events.move.out[i-1, ] <- protein.all.data.T$events.move
   
   #store the acceptance of protein
 acceptance.rate.protein[i-1]<- protein.all.data.T$acceptance.protein
 #each one parameter, store one data set 
 
 #store the ith interval protein birth number
 total.interval.protein.birth.number[i-1,]<- protein.all.data.T$interval.protein.birth.number
 #store all the data
 store.all.data.T[[i]]<- all.data.T
#store the total protein birth number
 protein.birth.number[i]<- length(which(all.data.T$jump.type == "birth.event"))
 
  }
  
  one.sample.mean<- colMeans(one.para.out)
  
   return(list(store.all.data.T=store.all.data.T, 
               one.para.out=one.para.out,
               one.sample.mean=one.sample.mean, 
               acceptance.rate.protein= acceptance.rate.protein,
               protein.birth.number= protein.birth.number,
               total.interval.protein.birth.number= total.interval.protein.birth.number ,
               total.events.move.out= total.events.move.out) )
}
  
```




```{r}

one.mcmc.run.simple.para.acceptance.fn<- function(all.data.T, partial.data.out, weight.prob, para.iter, initial.para.states, N, alp1,alp2, alp3, bet1, bet2,bet3, return.data.index ){
  
    #input the scaled theta, where theta2 is not theta2/N here
  
  one.para.out<- matrix(NA, nrow= (para.iter+1), ncol=3)
  one.para.out[1,]<- initial.para.states
  colnames( one.para.out)<- c("theta1", "theta2", "theta3")
  
  total.events.move.out<- as.data.frame( matrix(NA, nrow= para.iter, ncol=3) )
  colnames(total.events.move.out)<- c("add","shift","delete")
  
  #start each data, one row of parameter corresponding to one of data matrix
  store.all.data.T<- vector("list", (para.iter+1))
  store.all.data.T[[1]]<- all.data.T
  
  acceptance.rate.protein<- rep(NA,l=para.iter)
  protein.birth.number<- rep(NA,l=para.iter)
  #start from true protein birth number
  protein.birth.number[1]<- length(which(all.data.T$jump.type == "birth.event"))
  
  total.interval.protein.birth.number<- matrix(NA, nrow=para.iter, ncol= (nrow(partial.data.out)-1) )
  
  for(i in 2:(para.iter+1)){
#step 1: Gibbs of paramters 
#conditional on the [0,T] of full-observed data
#gibbs update: which is consider the full trajectory not in the ith interval!
 ##################################################################################################
 #this is the info for the summary of data to input the  the full-conditional of parameter 
full.data.summary.out.T.value<- cond.para.based.full.data.fn(all.data.T = all.data.T)
  

#this is the function of full conditional of parameter out given the full conditional data
  theta.all.can<- full.cond.theta.gamma.fn(full.data.summary.out.T= full.data.summary.out.T.value,
                        alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3, N=N)
  
  
   one.para.out[i,]<- theta.all.can
##################################################################################################
  
  
#step 2: Gibbs of mRNA
#conditional on the [0,T] of protein birth data & N jumps of the total reaction times of mRNA data 
#because the mRNA jumps is conditional on the protein birth events
#2.1 input of protein birth data(X2.data)
##################################################################################################
   
mrna.upated.all.data.T <- ful.cond.mrna.jumps.updated.fixed.na.fn(theta= theta.all.can, partial.data.out = partial.data.out, all.data.T = all.data.T,  N=N)
##################################################################################################
#step3: MH of protein  
#this is the update of all intervals 
##################################################################################################
   
 
protein.all.data.T <- protein.all.data.update.simple.para.acceptance.fn(all.data.T= mrna.upated.all.data.T, partial.data.out=partial.data.out, weight.prob=weight.prob, theta=theta.all.can, N=N )
  
   
   all.data.T<-  protein.all.data.T$updated.all.data.T  #update the current species data
   
   total.events.move.out[i-1, ] <- protein.all.data.T$events.move
   
   #store the acceptance of protein
 acceptance.rate.protein[i-1]<- protein.all.data.T$acceptance.protein
 #each one parameter, store one data set 
 
 #store the ith interval protein birth number
 #total.interval.protein.birth.number[i-1,]<- protein.all.data.T$interval.protein.birth.number
 #store all the data
 
 if( return.data.index == TRUE){
 store.all.data.T[[i]]<- all.data.T
 }
 
#store the total protein birth number
# protein.birth.number[i]<- length(which(all.data.T$jump.type == "birth.event"))
 
  }
  
#  one.sample.mean<- colMeans(one.para.out)
  
   return(list( store.all.data.T=store.all.data.T, 
               one.para.out=one.para.out,
               #one.sample.mean=one.sample.mean, 
               acceptance.rate.protein= acceptance.rate.protein,
              # protein.birth.number= protein.birth.number,
              # total.interval.protein.birth.number= total.interval.protein.birth.number ,
               total.events.move.out= total.events.move.out) )
}
  
```



#create a folder, save the time series plot of parameters and the data of species states
```{r}

plotProteinMRNA<-function(total.out, para.file, folder.file, data.file,data2.file, true.posterior.paras ,true.theta, N,k, protein.nb.data.file ,ith.interval.protein.data.file, theta.ratio.file){
  #create a folder
  if (!dir.exists(folder.file)) {
 dir.create(folder.file, recursive = TRUE, showWarnings = FALSE)
}

  
  
 para.file.name<-  file.path(folder.file, para.file)
 
  ## -------- open one PDF device ----------
  pdf(para.file.name)
  par(mfrow = c(3, 1))
      
  para.res<- total.out$one.para.out
  

ts.plot(para.res[,1],
          ylab = expression(theta[1]),  
        main = bquote("Partial:" ~ theta[1] ~
                      ",  N = " ~ .(N) ~
                      ",  chop number= " ~ 2^{.(k)})
        ,
        col = "black")
abline(h = true.theta[1], col = "blue", lwd = 2)
abline(h = true.posterior.paras[1], col = "red", lwd = 2)  
legend("topright",
       legend = c("true theta", "true posterior theta"),
       col    = c("blue", "red"),
       lty    = c(1, 1),
       lwd    = c(2, 2),
       bg     = "white" ,
       cex    = 0.7,        # ← text and line sizes scaled to 80 %
       seg.len = 2)         # ← (optional) shorten line segments


ts.plot(para.res[,2],
          ylab = expression(theta[2]), 
        main = bquote("Partial:" ~ theta[2] ~
                      ",  N = " ~ .(N) ~
                      ",  chop number= " ~ 2^{.(k)})
        ,col = "black")
abline(h = true.theta[2], col = "blue", lwd = 2)
abline(h = true.posterior.paras[2], col = "red", lwd = 2)    
legend("topright",
       legend = c("true theta", "true posterior theta"),
       col    = c("blue", "red"),
       lty    = c(1, 1),
       lwd    = c(2, 2),
       bg     = "white" ,
       cex    = 0.7,        # ← text and line sizes scaled to 80 %
       seg.len = 2)         # ← (optional) shorten line segments


ts.plot(para.res[,3],
          ylab = expression(theta[3]),main = bquote("Partial:" ~ theta[3] ~
                      ",  N = " ~ .(N) ~
                      ",  chop number= " ~ 2^{.(k)}),col = "black")
abline(h = true.theta[3], col = "blue", lwd = 2)
abline(h = true.posterior.paras[3], col = "red", lwd = 2)    
legend("topright",
       legend = c("true theta", "true posterior theta"),
       col    = c("blue", "red"),
       lty    = c(1, 1),
       lwd    = c(2, 2),
       bg     = "white" ,
       cex    = 0.7,        # ← text and line sizes scaled to 80 %
       seg.len = 2)         # ← (optional) shorten line segments

dev.off()
 cat("figure saved to", normalizePath(para.file.name), "\n")


######################################################################################################
#total species states
data.file.name<-  file.path(folder.file, data.file)
  
 pdf(data.file.name)

species.all.data<- total.out$store.all.data.T
 
## ------------------------------------------------------------------
## 0. convenience handles
## ------------------------------------------------------------------
true_dat <- species.all.data[[1]]          # the “truth”
sim_list <- species.all.data[-1]           # all the rest
#para.iter <- length(species.all.data)      # total iterations

## ------------------------------------------------------------------
## 1. determine axis limits
## ------------------------------------------------------------------
xlim <- range(true_dat$jump.times,
              unlist(lapply(sim_list, `[[`, "jump.times")))

ylim <- range(true_dat$SpeciesProtein,
              true_dat$SpeciesmRNA,
              unlist(lapply(sim_list, `[[`, "SpeciesProtein")),
              unlist(lapply(sim_list, `[[`, "SpeciesmRNA")))

## ------------------------------------------------------------------
## 2. blank plot
## ------------------------------------------------------------------
plot(NULL, type = "n",
     xlim = xlim, ylim = ylim,
     xlab = "time", ylab = "count",
      main = bquote("Protein (solid) & mRNA (dashed)"
                   ~ ",  N = " ~ .(N)
                   ~ ",  chop number= " ~ 2^{.(k)}))

## ------------------------------------------------------------------
## 3. add simulated iterations (grey, semi-transparent)
## ------------------------------------------------------------------
cols       <- rainbow(para.iter)                  # vivid rainbow hues
cols.alpha <- adjustcolor(cols, alpha.f = 0.70)   # 70 % opacity
 

for (i in seq_along(sim_list)) {
  dat <- sim_list[[i]]
  col.i <- cols.alpha[i]

  ## Protein – solid, translucent rainbow
  lines(dat$jump.times, dat$SpeciesProtein,
        type = "s", col = col.i, lwd = 0.5)

  ## mRNA – dashed, same colour
  lines(dat$jump.times, dat$SpeciesmRNA,
        type = "s", col = col.i, lwd = 0.5, lty = 2)
}

## ------------------------------------------------------------------
## 4. add the true trajectory (black)
## ------------------------------------------------------------------
lines(true_dat$jump.times, true_dat$SpeciesProtein,
      type = "s", col = "black", lwd = 1)

lines(true_dat$jump.times, true_dat$SpeciesmRNA,
      type = "s", col = "black", lwd = 1)

## ------------------------------------------------------------------
## 5. vertical lines for Partial.obse.times in the true data
## ------------------------------------------------------------------
v_times <- species.all.data[[2]]$jump.times[ species.all.data[[2]]$jump.type == "Partial.obse.times" ]
abline(v = v_times, col = "red", lty = 3, lwd = 1)

## ------------------------------------------------------------------
## 6. legend
## ------------------------------------------------------------------
legend("topright",
       legend = c("true Protein", "true mRNA", "Partial.obse.times"),
       col    = c("black", "black", "red"),
       lty    = c(1, 2, 3),
       lwd    = c(2, 2, 1),
       bg     = "white" ,
       cex    = 0.8,        # ← text and line sizes scaled to 80 %
       seg.len = 2)         # ← (optional) shorten line segments

dev.off()
 
 cat("figure saved to", normalizePath(data.file.name), "\n")
#######################################################################################################
 #only protein states
######################################################################################################
data2.file.name <- file.path(folder.file, data2.file)
pdf(data2.file.name)

species.all.data <- total.out$store.all.data.T

## ------------------------------------------------------------------
## 0. convenience handles
## ------------------------------------------------------------------
true_dat <- species.all.data[[1]]          # the “truth”
sim_list <- species.all.data[-1]           # all the rest
para.iter <- length(species.all.data)      # total iterations

## ------------------------------------------------------------------
## 1. determine axis limits  (Protein only)
## ------------------------------------------------------------------
xlim <- range(true_dat$jump.times,
              unlist(lapply(sim_list, `[[`, "jump.times")))

ylim <- range(true_dat$SpeciesProtein,
              unlist(lapply(sim_list, `[[`, "SpeciesProtein")))

## ------------------------------------------------------------------
## 2. blank plot
## ------------------------------------------------------------------
plot(NULL, type = "n",
     xlim = xlim, ylim = ylim,
     xlab = "time", ylab = "Protein count",
     main = bquote("Protein trajectories"
                   ~ ",  N = " ~ .(N)
                   ~ ",  chop number = " ~ 2^{.(k)}))

## ------------------------------------------------------------------
## 3. add simulated iterations (grey, semi-transparent)
## ------------------------------------------------------------------
cols       <- rainbow(para.iter)                  # vivid rainbow hues
cols.alpha <- adjustcolor(cols, alpha.f = 0.70)   # 70 % opacity

for (i in seq_along(sim_list)) {
  dat   <- sim_list[[i]]
  col.i <- cols.alpha[i]

  ## Protein – solid, translucent rainbow
  lines(dat$jump.times, dat$SpeciesProtein,
        type = "s", col = col.i, lwd = 0.5)
}

## ------------------------------------------------------------------
## 4. add the true trajectory (black)
## ------------------------------------------------------------------
lines(true_dat$jump.times, true_dat$SpeciesProtein,
      type = "s", col = "black", lwd = 0.3)

## ------------------------------------------------------------------
## 5. vertical lines for Partial.obse.times in the true data
## ------------------------------------------------------------------
v_times <- species.all.data[[2]]$jump.times[
             species.all.data[[2]]$jump.type == "Partial.obse.times"
           ]
abline(v = v_times, col = "red", lty = 3, lwd = 1)

## ------------------------------------------------------------------
## 6. legend
## ------------------------------------------------------------------
legend("topright",
       legend = c("true Protein", "Partial.obse.times"),
       col    = c("black", "red"),
       lty    = c(1, 3),
       lwd    = c(2, 1),
       bg     = "white",
       cex    = 0.8,
       seg.len = 2)

dev.off()

cat("figure saved to", normalizePath(data2.file.name), "\n")
###################################################################################################
###################################################################################################
######################################################################################################
######

protein.nb.data.file.name<-  file.path(folder.file, protein.nb.data.file)
pdf(protein.nb.data.file.name)

protein.nb.data<- total.out$protein.birth.number

# set up a step plot
plot(1:length(protein.nb.data),                  # x-axis: 1, 2, 3, 4
     protein.nb.data,                           # y-axis values
     type = "s",                                # "s" = stair-step
     pch  = 16,                                 # solid dots on each step
     xlab = "Step index",
     ylab = "Protein birth count",
     main = "Protein births across steps")

# emphasise the points themselves (optional but tidy)
points(1:length(protein.nb.data), protein.nb.data, pch = 16,cex=0.3)

# add a horizontal reference line at the first value
abline(h = protein.nb.data[1], col = "blue", lty = 2, lwd = 2)

# label that reference line
text(x     = 0.8*length(protein.nb.data),           # far right of the plot
     y     = protein.nb.data[1], 
     labels = "true birth number", 
     pos   = 3,  col = "blue")                  # pos = 3 puts it above the line

 dev.off()
 
 cat("figure saved to", normalizePath(protein.nb.data.file.name), "\n")

 
 
######################################################################################################
######################################################################################################
 
ith.interval.protein.data.file.name<-  file.path(folder.file, ith.interval.protein.data.file )
pdf(ith.interval.protein.data.file.name)

 ## -----------------------------------------------------------
## 0.  Data
## -----------------------------------------------------------
mat <- total.out$total.interval.protein.birth.number        # nIter × 9 matrix
nIter <- nrow(mat)
x <- seq_len(ncol(mat))                           # 1,2,…,9

## -----------------------------------------------------------
## 1.  Base plot: all iterations in light grey
## -----------------------------------------------------------
matplot(x, t(mat), type = "l", lty = 1, col = "grey80",
        xlab = "Interval (column index)",
        ylab = "Protein-birth count",
        main = "All iterations (rows) across intervals")

# 2. Highlight first (row 1) and last (row nIter)
special.rows <- c(1, nIter)
special.col  <- c("royalblue", "firebrick")

matlines(x, t(mat[special.rows, ]), lty = 1, lwd = 2, col = special.col)

# Optional legend
legend("topright",
       legend = c("all other iterations", "first iteration", "last iteration"),
       col    = c("grey80", special.col),
       lwd    = c(1, 2, 2),
       bty    = "n", cex = 0.75)
dev.off()
 
 cat("figure saved to", normalizePath(ith.interval.protein.data.file.name), "\n")


 
 ######################################################################################################
theta.ratio.file.name<-  file.path(folder.file, theta.ratio.file)
  
pdf(theta.ratio.file.name)
 
  para.res<- total.out$one.para.out
  
   theta.ratio.data<- para.res[,2]/para.res[,3]
   
 plot(theta.ratio.data, 
      type = "l",
     main  = bquote(
  "Partial:" ~~ theta[2]/theta[3] *
  ",  N = "  ~ .(N) *
  ",  chop number = " ~ 2^{.(k)}
),
     xlab  = "Iteration",
     ylab  = expression(theta[2] / theta[3]),
     col   = "steelblue")
 
abline(h = mean(theta.ratio.data), col = "firebrick", lwd = 2)
abline(h = true.theta[2] / true.theta[3], col = "darkgreen", lwd = 2, lty = 3)
legend("topright",
       legend = c("MCMC theta ratio", "Posterior mean", "True ratio"),
       col    = c("steelblue",  "firebrick",     "darkgreen"),
       lty    = c(1,            2,              3),
       lwd    = c(1,            2,              2),
       bty    = "n",            # no border
       cex    = 0.9)
  
dev.off()
 
 cat("figure saved to", normalizePath(theta.ratio.file.name), "\n")


 ######################################################################################################
#accepance.file.name<-  file.path(folder.file, acceptance.file)
  
# pdf(accepance.file.name)

#protein.acceptance.rates<- total.out$acceptance.rate.protein
 
# plot(protein.acceptance.rates,
#     type = "l",          # “l” = lines
#     xlab = "Iteration",
#     ylab = "Acceptance rate",
#     main = "MCMC acceptance rate over iterations",
#     ylim = c(0, 1))      # acceptance rate is a probability
#abline(h = mean(protein.acceptance.rates), col = "red", lwd = 2, lty = 2)  # overall mean
#legend("topright", legend = sprintf("mean = %.3f", mean(protein.acceptance.rates)),
#       col = "red", lty = 2, lwd = 2, box.lwd = 0, cex = 0.8)

#dev.off()
 
# cat("figure saved to", normalizePath(accepance.file.name), "\n")

}


```



#fn:mcmc given one data
```{r}
 
fixed.one.data.mcmc.run.fn<- function(true.theta, chop.number.times.region, same.data.index, N, data.iter, para.iter,weight.prob,know.reaction.final.index, T.protein.unknown, ith.iter ){
  
alp1=0.0001
alp2=0.0001
bet1=0.0001
bet2=0.0001
alp3=0.0001
bet3=0.0001

  theta1=true.theta[1]     #true parameter
theta2.o= true.theta[2]
theta3= true.theta[3]
  
S0 <- c(N, 0)
theta.ori<-c(theta1,theta2.o/N,theta3)

#fixed one data
one.par.mean.out<- matrix(NA, nrow=data.iter, ncol=3)
#true.posterior.paras<- one.par.mean.out
store.par.mean.out.chops<- vector("list",length(chop.number.times.region))
theta.ratio.mean.out<- rep(NA, l=length(chop.number.times.region))


  for(it in 1:length(chop.number.times.region)){
    store.par.mean.out.chops[[it]]<- one.par.mean.out
  }
 

  #start data
for( j in 1:data.iter ){
  
if( same.data.index == FALSE){
result.out<- ssa.f.total(theta=theta.ori, S0=S0)

  
}else{ 
  set.seed(123)
    result.out <- ssa.f.total(theta=theta.ori, S0=S0)
    rm(.Random.seed, envir = .GlobalEnv)   # back to ordinary RNG
    }
 
################################################################   
  
 
  for(k in 1:length(chop.number.times.region)){
  
################################################################   
  data.p.out <- chop.time.equally.fn(all.data = result.out, chop.number.times=chop.number.times.region[k],  know.reaction.final.index=know.reaction.final.index, T.protein.unknown=T.protein.unknown )
  #after chop, out put the partial time data and the true full observed data
  partial.data.out<- data.p.out$partial.all.species.data  
  #this is the partial time observation
  #start true of all species states  and parameter
  all.data.T<- data.p.out$with.T.all.data 
 #############################################################  
   #this is true posterior parameter

 true.data.T.summary.out<- cond.para.based.full.data.fn(all.data.T =all.data.T )
  true.posterior.paras<- full.cond.theta.gamma.fn(full.data.summary.out.T = true.data.T.summary.out,
                        alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3, N=N)
  
 

  start <- Sys.time()
  
  
  one.mcmc.out<- one.mcmc.run.fn(para.iter= para.iter, ith.iter= ith.iter,
                        all.data.T = all.data.T, partial.data.out=partial.data.out, 
                          initial.para.states= true.theta,
                        weight.prob=weight.prob, N=N,
                   alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3)
  
  end <- Sys.time()
  total.time <- difftime(end, start, units = "auto")
  cat("For data iter",j, "simulation time is", format(total.time), "\n","with N=", N)
  
  
  #save the sample mean
   store.par.mean.out.chops[[k]][j,]<-  one.mcmc.out$one.sample.mean
   
   theta.ratio.mean.out[k]<- mean( one.mcmc.out$one.para.out[,2]/one.mcmc.out$one.para.out[,3] )
     
   print(paste("finish chop", chop.number.times.region[k], "data iter", j ))


  
 ####################################################################################################  
 #save the plots to a folder  
  para.file <- paste0("para.out-chop.times=",chop.number.times.region[k] , ".pdf")
####################################################################################################  
file_name <-  para.file
write.csv(one.mcmc.out$one.para.out, file = file_name, row.names = FALSE)
cat("CSV saved to", normalizePath(file_name), "\n")
####################################################################################################  

   
   data.file<- paste0("data.out-chop.times=", chop.number.times.region[k], ".pdf")
    data2.file<- paste0("Protein.out-chop.times=", chop.number.times.region[k], ".pdf")
   acceptance.file<- paste0("acceptance.out-chop.times=", chop.number.times.region[k], ".pdf")
   protein.nb.data.file<- paste0("protein.nb.data-chop.times=", chop.number.times.region[k], ".pdf")
     ith.interval.protein.data.file<-paste0("ith.interval.protein.data-chop.times", chop.number.times.region[k], ".pdf")
       
theta.ratio.file<- paste0("theta.ratio-chop.times=", chop.number.times.region[k], ".pdf")
  
      ## choose your day-month pattern once
today_tag <- format(Sys.time(), "%d%b-%H")     # "18Jun-1123" (adds hour+min)

## build the full name inside your loop
folder.file <- paste0(today_tag, "-ith-iter-with-Data", j, ", N=", N)


  plotProteinMRNA(total.out =  one.mcmc.out, para.file = para.file, 
                  data.file =data.file,data2.file=data2.file,
                  folder.file=folder.file, 
                  true.posterior.paras=true.posterior.paras, true.theta = true.theta , N=N , k=chop.number.times.region[k], 
                  acceptance.file =acceptance.file,
                  protein.nb.data.file= protein.nb.data.file, 
                  ith.interval.protein.data.file= ith.interval.protein.data.file,
                  theta.ratio.file= theta.ratio.file)
  
  
  
  }
  
}
  return(list(store.par.mean.out.chops=store.par.mean.out.chops,true.posterior.paras=true.posterior.paras, theta.ratio.mean.out=theta.ratio.mean.out ))
  
}
```

#####################################################################################

#start here
```{r}
true.theta=  c(0.5,90,1)
chop.number.times.region<- c(3,4,5,6)
same.data.index= TRUE
data.iter=1
para.iter= 1000
weight.prob<- c(1/3,1/3,1/3)
N=100
ith.iter= 1


total.out<- fixed.one.data.mcmc.run.fn(true.theta = true.theta,
    chop.number.times.region=chop.number.times.region, same.data.index=same.data.index,N=N, 
                  data.iter=data.iter,  para.iter=para.iter,
    ith.iter=ith.iter,
    weight.prob = weight.prob, T.protein.unknown=10, 
    know.reaction.final.index = FALSE
    ) 
```



#this is the one in Rscript, draw the parameter with given dataset


```{r}
given.one.data.mcmc.run.simple.fn<- function( given.result.out, chop.number.times.region, N, true.theta,start.theta,  para.iter,weight.prob, know.reaction.final.index, T.protein.unknown ){
  
  alp1=0.0001
  alp2=0.0001
  bet1=0.0001
  bet2=0.0001
  alp3=0.0001
  bet3=0.0001
  
  
  #fixed one data
  #one.par.mean.out<- matrix(NA, nrow=1, ncol=3)
  #true.posterior.paras<- one.par.mean.out
  #store.par.mean.out.chops<- vector("list",length(chop.number.times.region))
  #theta.ratio.mean.out<- rep(NA, l=length(chop.number.times.region))
  
  
  # for(it in 1:length(chop.number.times.region)){
  #   store.par.mean.out.chops[[it]]<- one.par.mean.out
  # }
  
  
  #start data
  
  for(k in 1:length(chop.number.times.region)){
    
    ################################################################   
    data.p.out <- chop.time.equally.fn(all.data = given.result.out, chop.number.times=chop.number.times.region[k],  know.reaction.final.index=know.reaction.final.index, T.protein.unknown=T.protein.unknown )
    #after chop, out put the partial time data and the true full observed data
    partial.data.out<- data.p.out$partial.all.species.data  
    #this is the partial time observation
    #start true of all species states  and parameter
    all.data.T<- data.p.out$with.T.all.data 
    #############################################################  
    #this is true posterior parameter
    
    true.data.T.summary.out<- cond.para.based.full.data.fn(all.data.T =all.data.T )
    true.posterior.paras<- full.cond.theta.gamma.fn(full.data.summary.out.T = true.data.T.summary.out,
                                                    alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3, N=N)
    
    
    
    start <- Sys.time()
    
    
    one.mcmc.out<- one.mcmc.run.simple.para.acceptance.fn(para.iter= para.iter,
                                                          all.data.T = all.data.T, partial.data.out=partial.data.out, 
                                                          initial.para.states= start.theta,
                                                          weight.prob=weight.prob, N=N,
                                                          alp1=alp1, alp2=alp2, bet1 = bet1, bet2 = bet2,alp3=alp3, bet3=bet3, return.data.index = FALSE)
    
    end <- Sys.time()
    total.time <- difftime(end, start, units = "auto")
    cat("For simulation time is", format(total.time), "\n","with N=", N)
    
    
    #save the sample mean
    #store.par.mean.out.chops[[k]][j,]<-  one.mcmc.out$one.sample.mean
    
    # theta.ratio.mean.out[k]<- mean( one.mcmc.out$one.para.out[,2]/one.mcmc.out$one.para.out[,3] )
    
    print(paste("finish chop", chop.number.times.region[k], "one data iter" ))
    
    ####################################################################################################  
    #save the plots to a folder  
    para.file <- paste0("para.out-chop.times=",chop.number.times.region[k] , ",N=", N, ".csv")
    write.csv(one.mcmc.out$one.para.out, file = para.file , row.names = FALSE)
    cat("CSV saved to", normalizePath(para.file ), "\n")
    ####################################################################################################  
    acceptance.file<- paste0("acceptance.out-chop.times=", chop.number.times.region[k], ",N=", N, ".csv")
    write.csv(  one.mcmc.out$acceptance.rate.protein , file = acceptance.file , row.names = FALSE)
    cat("CSV saved to", normalizePath(acceptance.file ), "\n")
    ####################################################################################################  
    events.move.file<- paste0("events-move-chop.times=", chop.number.times.region[k], ",N=", N, ".csv")
    write.csv(one.mcmc.out$total.events.move.out, file =  events.move.file , row.names = FALSE)
    cat("CSV saved to", normalizePath( events.move.file ), "\n")
    ####################################################################################################  
    #  data.file<- paste0("data.out-chop.times=", chop.number.times.region[k], ".pdf")
    #   data2.file<- paste0("Protein.out-chop.times=", chop.number.times.region[k], ".pdf")
    #acceptance.file<- paste0("acceptance.out-chop.times=", chop.number.times.region[k], ".pdf")
    #  protein.nb.data.file<- paste0("protein.nb.data-chop.times=", chop.number.times.region[k], ".pdf")
    #   ith.interval.protein.data.file<-paste0("ith.interval.protein.data-chop.times", chop.number.times.region[k], ".pdf")
    
    #theta.ratio.file<- paste0("theta.ratio-chop.times=", chop.number.times.region[k], ".pdf")
    
    ## choose your day-month pattern once
    #today_tag <- format(Sys.time(), "%d%b-%H")     # "18Jun-1123" (adds hour+min)
    
    ## build the full name inside your loop
    #folder.file <- paste0(today_tag, "-One-Data, N=", N)
    
    
    # plotProteinMRNA(total.out =  one.mcmc.out, para.file = para.file, 
    #                 data.file =data.file,data2.file=data2.file,
    #                  folder.file=folder.file, 
    #                  true.posterior.paras=true.posterior.paras, true.theta = true.theta , N=N , k=chop.number.times.region[k], 
    #                  #acceptance.file =acceptance.file,
    #                  protein.nb.data.file= protein.nb.data.file, 
    #                  ith.interval.protein.data.file= ith.interval.protein.data.file,
    #                  theta.ratio.file= theta.ratio.file)
    
    
    
  }
  
  
  # return(list(#store.par.mean.out.chops=store.par.mean.out.chops,
  #   true.posterior.paras=true.posterior.paras, 
  #   theta.ratio.mean.out=theta.ratio.mean.out ))
  
}

```






#save the theta ratio



```{r}
##########################################################################################################
## ------------------------------------------------------------------
## 0.  Inputs assumed to exist
## ------------------------------------------------------------------
theta_ratio_mean <- total.out$theta.ratio.mean.out     # numeric vector
chop.number      <- chop.number.times.region           # same length vector

## sanity check
stopifnot(length(theta_ratio_mean) == length(chop.number))

## ------------------------------------------------------------------
## 1.  Build a data frame (LONG format: one row per region)
## ------------------------------------------------------------------
out_df <- data.frame(
  chop.number.times.region = chop.number,
  theta23.mean             = theta_ratio_mean
)

## ------------------------------------------------------------------
## 2.  Write to CSV (file name includes today’s date for reproducibility)
## ------------------------------------------------------------------
file_name <- sprintf("theta23_mean_by_chop_%s.csv", format(Sys.Date(), "%Y%m%d"))
write.csv(out_df, file = file_name, row.names = FALSE)

cat("CSV saved to", normalizePath(file_name), "\n")

##########################################################################################################
#
## ------------------------------------------------------------------
## 2.  Plot to a PDF file
## ------------------------------------------------------------------
pdf_name <- sprintf("theta23_mean_plot_%s.pdf",
                    format(Sys.Date(), "%Y%m%d"))
pdf(pdf_name, width = 6, height = 4)
true_ratio        <- true.theta[2] / true.theta[3]
y_range_raw       <- range(theta_ratio_mean, true_ratio, na.rm = TRUE)
pad               <- 0.05 * diff(y_range_raw)          # 5 % head- and foot-room
y_range           <- y_range_raw + c(-pad, pad)


plot(chop.number, theta_ratio_mean,
     type = "b", pch = 19, lwd = 2, col = "steelblue",
     xlab = "chop.number.times.region",
     ylab = expression("sample mean of" ~~ theta[2]/theta[3]),
     main = expression("Posterior means of" ~~ theta[2]/theta[3]
                       ~~ "across chop regions"),
     ylim = y_range )
grid()

## --- horizontal line at the true ratio ----
true_ratio <- true.theta[2] / true.theta[3]      # assumes true.theta vector exists
abline(h = true_ratio, col = "firebrick", lty = 2, lwd = 2)

## --- legend ----

par(xpd = NA)   #allow drawing outside the plot

## add the legend just above the top edge
legend("top",
       inset = c(0, -0.15),        # 5 % of plot‐height ABOVE the frame
       horiz = TRUE,
       legend = c("MCMC mean", "true ratio"),
       col    = c("steelblue", "firebrick"),
       lty    = c(1, 2),
       pch    = c(19, NA),
       lwd    = c(2, 2),
       bty    = "n")               # no border box

## restore normal clipping (optional but good practice)
par(xpd = FALSE)
dev.off()
cat("PDF figure saved to", normalizePath(pdf_name), "\n")
```



#with data iterations, plot the sample means
###############################################################################
```{r, include=FALSE}
 
#save the sample mean
folder <- sprintf("csv_chops_N=%d", N)   # "csv_chops_N=150"
if (!dir.exists(folder)) dir.create(folder)

for (k in seq_along(out$store.par.mean.out.chops)) {
  m <- out$store.par.mean.out.chops[[k]]
  fn <- sprintf("chop_%02d.csv", k)
  write.csv(m,
            file = file.path(folder, fn),
            row.names = FALSE)
}


 

plotChopHistos <- function(out,true.theta,N,
                           pdf.file = "theta_hists.pdf",
                           height.per = 2.4,   # inches per row    (PDF device)
                           width.per  = 3.0,   # inches per column (PDF device)
                           breaks = "FD") {
  
  chops.list <- out$store.par.mean.out.chops   # list of matrices
  R <- length(chops.list)                      # rows  (chop regions)
  C <- ncol(chops.list[[1]])                   # cols  (parameters)
  param.names <- paste0("theta", seq_len(C))
  
  ## -------- open PDF device -------------------------------------------
  pdf(pdf.file,
      width  = width.per  * C,
      height = height.per * R)
  on.exit(dev.off(), add = TRUE)               # close on exit
  
  ## -------- layout: R rows × C columns --------------------------------
  par(mfrow = c(R, C),
      mar   = c(3, 4, 2, 1),                  # inner margins
      oma   = c(4, 0, 3, 0))                  # outer margin for titles
  
 
  ## -------- loop over chops & parameters ------------------------------
  for (i in seq_len(R)) {
    dat.mat <- chops.list[[i]]
    
    
    for (j in seq_len(C)) {
      
      x <- dat.mat[ , j]                      # vector of means
      
       main.lab <- sprintf("chop %d – %s, N = %d",
                i,               # %d → chop index
                param.names[j],  # %s → parameter name
                N)               # %d → N
  
      hist(x, breaks = breaks, freq = TRUE,
           main = main.lab,
           xlab = "", ylab = "freq",
           col = "grey85", border = "grey50")
      
      ## blue: true theta
      abline(v = true.theta[j],
             col = "blue", lwd = 2)
      
      ## red: posterior theta
      abline(v = out$true.posterior.paras[j],
             col = "red", lwd = 2)
      
      ## add legend only in the first panel
      if (i == 1 && j == 1) {
        legend("topright",
               legend = c("true theta", "posterior theta"),
               col    = c("blue",       "red"),
               lty    = 1,
               lwd    = 2,
               bg     = "white",
               cex    = 0.7,
               seg.len = 2)
      }
    }
  }
  
  ## -------- shared axis label -----------------------------------------
  mtext("value", side = 1, outer = TRUE, line = 2.0)
  
  invisible(normalizePath(pdf.file))
}

outfile <- plotChopHistos(out= out,true.theta= true.theta,N=N,
                          pdf.file  = " para.out-chop.times=1.pdf",
                          breaks    = 25)   # any hist() argument goes here
cat("figure saved to", outfile, "\n")

```


#Prior 
#Plot-prior-vs-posterior-density
```{r}
#give one data
true.theta<- c(0.5,90,3)
  theta1=true.theta[1]     #true parameter
theta2.o= true.theta[2]
theta3= true.theta[3]
 N=100
S0 <- c(N, 0)
theta.ori<-c(theta1,theta2.o/N,theta3)
result.out<- ssa.f.total(theta=theta.ori, S0=S0)

################################################################   
know.reaction.final.index= FALSE
T.protein.unknown= 10
  data.p.out <- chop.time.equally.fn(all.data = result.out, chop.number.times=3,  know.reaction.final.index=know.reaction.final.index, T.protein.unknown=T.protein.unknown )
  #after chop, out put the partial time data and the true full observed data
  partial.data.out<- data.p.out$partial.all.species.data  
  #this is the partial time observation
  #start true of all species states  and parameter
  all.data.T<- data.p.out$with.T.all.data 
 #############################################################  
  
full.data.summary.out.T.value<- cond.para.based.full.data.fn(all.data.T = all.data.T)
  

#this is the function of full conditional of parameter out given the full conditional data
na= full.data.summary.out.T.value$na
nb= full.data.summary.out.T.value$nb
nd= full.data.summary.out.T.value$nd
int.X1= full.data.summary.out.T.value$int.X1
int.X2= full.data.summary.out.T.value$int.X2

x1.seq<- seq(true.theta[1]*0.1, true.theta[1]*2,l=100)
x2.seq<- seq(true.theta[2]*0.1, true.theta[2]*2, l=100)
x3.seq<- seq(true.theta[3]*0.1, true.theta[3]*2,l=100)
```

```{r}
priors <- list(
  "Uninformative" = list(
    alp = c(1e-4, 1e-4, 1e-4),
    bet = c(1e-4, 1e-4, 1e-4)
  ),
  "Medium-informative" = list(
    alp = c(2, 2, 2),
    bet = 2 /(0.5* true.theta )        # vectorised: 2/θ₁, 2/θ₂, 2/θ₃
  ),
  
  "Strong-Informative"=list(
    alp = c(2, 2, 2),
    bet = 2/(1*true.theta)           
  )
)
```


```{r}
## -----------------------------------------------------------
## 3.  PLOTTING
## -----------------------------------------------------------
par(mfrow = c(length(priors),3),      # 3 rows (θ₁, θ₂, θ₃) × 2 cols (priors)
    mar   = c(4, 4, 2.5, 1),           # room for column titles
    las   = 1)

theta.seq.list <- list(x1.seq, x2.seq, x3.seq)   # grid for each θ
suff.n  <- c(na, nb, nd)                         # sample size for each θ
suff.int <- c(int.X1, int.X1 / N, int.X2)        # ∑ x for θ₁, scaled θ₂, θ₃

for (p.name in names(priors)) {
  pr <- priors[[p.name]]

  for (k in 1:3) {                       # k = 1,2,3 for θ₁, θ₂, θ₃
    shape <- suff.n[k]  + pr$alp[k]
    rate  <- suff.int[k] + pr$bet[k]
    dens  <- dgamma(theta.seq.list[[k]], shape = shape, rate = rate)

    plot(theta.seq.list[[k]], dens, type = "l",
         main = bquote(theta[.(k)]), xlab = "", ylab = "Posterior density")
    abline(v = true.theta[k], col = "blue", lwd = 1.3)

    ## add column heading above the top plot
    if (k == 1) {
      mtext(p.name, side = 3, line = 2, cex = 0.9, font = 2)
    }
  }
}

```
